A Reduced Order Model for Porous Media Flows: Coupling Linear Regression with Principal Component Analysis (PCA) using High-Order Singular Value Decomposition (HOSVD) in Upscaling Calculations for Permeability in Reservoir Simulations

### 2018-12-03 Papers/AppliedMaths2018/Resubmission/Stochastic_Upscaling_UK_AMM_2018b.tex
files to be modified in the next 250min

\subsubsection{Understanding the principal component spaces}\label{subsubsection:visualization_pcspaces}
Essentially, the existing permeability space is transformed (with SVD and HOSVD for 2- and 3-D permeability data, respectively) to principal component spaces (PCS). Within the PCS, traditional PCA dimensionality reduction can be optionally performed (\ie{feature selection or elimination}, see Section~\ref{subsubsection:pca_dimred_linreg} for more details), before interpolation is used to implement further data reduction in the principal component spaces. The reduced data in the PCS is then transformed back into the permeability space to obtain a reduced permeability space. An algorithm which impliments the procedure described here, is presented in Section~\ref{subsubsection:svdcase_preprocess_algorithm}

The PCS is a bit challenging to visualize because it represents projections which are embedded in the SVD/HOSVD factors of the original permeability field. An aid for visualizing the PCS for a 2D permeability field is to write out the $\mathbf{U}$, $\mathbf{\Sigma}$ and $\mathbf{V^{\intercal}}$ matrices (or SVD factors) and consider the singular vectors (\ie columns in $\mathbf{U}$ and rows in $\mathbf{V^{\intercal}}$) as separate PCS. A similar intuition can be extended to HOSVD for 3D permeability fields. The decomposition of a matrix into its factors using SVD  \cite{Cline_2007, Tharwat_2016} or HOSVD \cite{Kolda_2009, Lathauwer_2000} is covered in details in several published materials. However, for completeness, the appendix contains a brief introduction to SVD (\ref{subsection:svd_brief}) and HOSVD (\ref{subsection:hosvd_brief}).

>>>>

\subsubsection{Understanding the principal component spaces}\label{subsubsection:visualization_pcspaces}
Essentially, the existing permeability space is transformed (with SVD and HOSVD for 2- and 3-D permeability data, respectively) to principal component spaces (PCS). Within the PCS, traditional PCA dimensionality reduction can be optionally performed (\ie{feature selection or elimination}, Section~\ref{subsubsection:pca_dimred_linreg}), before interpolation is used to implement further data reduction in PCS. The reduced data in the PCS is then transformed back into the permeability space to obtain a reduced permeability space. An algorithm which impliments the procedure described here is presented in Section~\ref{subsubsection:svdcase_preprocess_algorithm}

The PCS is a slightly challenging to visualize because it represents projections which are embedded in the SVD/HOSVD factors of the original permeability field. An aid for visualizing the PCS for a 2D permeability field is to write out the $\mathbf{U}$, $\mathbf{\Sigma}$ and $\mathbf{V^{\intercal}}$ matrices (or SVD factors) and consider the singular vectors (\ie columns in $\mathbf{U}$ and rows in $\mathbf{V^{\intercal}}$) as separate PCS. A similar intuition can be extended to HOSVD for 3D permeability fields. The decomposition of a matrix into its factors using SVD  \cite{Cline_2007, Tharwat_2016} or HOSVD \cite{Kolda_2009, Lathauwer_2000} is covered in details in several published materials. However, for completeness, the appendix contains a brief introduction to SVD (\ref{subsection:svd_brief}) and HOSVD (\ref{subsection:hosvd_brief}).
